A fast algorithm for finding efficient decisions in multi-objective problems with black-box multiextremal functions

### 2022-01-06 2022/OM&S/paper.tex
Section 3

%\usepackage{color}
\usepackage{hyperref}

@@

For solving the problems of the family a global optimization technique based on reducing the multidimensional problem (\ref{conv_problem}) to an equivalent univariate subproblem [Strongin+Sergeyev, Pintér, Zhigljavsky+, Sergeyev+ Strongin, Paulavičius+, Sergeyev+ Kvasov]  in combination with one-dimensional information algorithm with accelerated convergence.

@@

>>>>

\usepackage{color}
\usepackage{hyperref}

@@

For solving the problems of the family a global optimization technique based on reducing the multidimensional problem (\ref{conv_problem}) to an equivalent univariate subproblem \cite{Strongin2000,Pinter1996,Zhigljavsky2008,Sergeyev2013,PaulaviciusZilinskas2014,Sergeyev2017} in combination with one-dimensional information algorithm with accelerated convergence.

@@

According to ARL applied to solving the problem (\ref{problem_fi}) its first $l \geq 2$  trials are executed at arbitrary points $t^1, \dots, t^l$ of the interval $[\alpha, \beta]$ including points $\alpha$ and $\beta$, and the function values $v^1, \dots, v^l$ are evaluated at these points, i.e., $v^s=\varphi(t^s)$, $1 \leq j \leq l$.

In order to choose a point $t^{n+1}$ of a new $(n+1)$-th trial after completing $n \geq l$ trials it is necessary to implement the following actions.

\begin{enumerate}
\item 
The points $t^1, \dots, t^n$ of completed trials and boundary points of the search interval $[\alpha, \beta]$ are ordered in increasing order and renumbered by subscripts, i.e.,
\begin{equation}\label{eq:15}
t_1 = \alpha < t_1 < \dots < t_{n-1} < t_n = \beta.
\end{equation}
The values $v_s = \varphi(t_s)$ are juxtaposed to the points  $t_s$, $1 \leq s \leq n$.
\item 
A real values $R(q)$ are assigned to all the intervals $(t_{q-1}, t_q)$, $2 \leq q \leq n$, formed by neighboring points from (\ref{eq:15}). $R(q)$ is called \textit{the characteristic} of the interval $(t_{q-1}, t_q)$ and is computed in accordance with the following rule.

\textcolor[rgb]{1,0,0}{If number of completed trials is not evenly divisible by $J$, where \textit{global-to-local ratio} $J$ is a predefined number, then }
\begin{equation}\label{eq:16}
R(q) = B(q),
\end{equation}
otherwise
\begin{equation}\label{eq:17}
R(q) = B(q)\left( \frac {\sqrt{(v_q - \varphi^*(n))(v_{q-1} - \varphi^*(n))}}{M} +(1.5)^{-\sigma}\right)^{-1}.
\end{equation}
Here
\begin{gather}\label{eq:18}
B(q) = (t_s - t_{s-1})^{1/N} + \frac{(v_q - v_{q-1})^2} {m^2 (t_s - t_{s-1})^{1/N}} -2\frac{v_{q-1} + v_q}{m},\\
\varphi^*(n) = \min_{1\leq j \leq n} {\varphi(t^j)}
\end{gather}
is minimal computed value of the objective function $\varphi(t)$ at trial points,
\begin{equation}\label{eq:19}
M = \max_{2 \leq  s \leq n} { \frac{|v_s - v_{s-1}|}{(t_s - t_{s-1})^{1/N}} }
\end{equation}
is underestimation of the H{\"o}lder constant obtained according to the trial results,
\begin{equation}\label{eq:20}
m = 
\begin{cases}
\begin{matrix}
rM, & M>0, \\
1, & M= 0,
\end{matrix}
\end{cases}
\end{equation}
is improved estimation of the H{\"o}lder constant due to method's parameter $r>1$, $\sigma$ is some integer parameter that influences the speed of convergence to the global minimum.

\item
Among all intervals, the interval   with the highest characteristic is selected
\begin{equation}\label{eq:21}
R(p) = \max_{2 \leq s \leq n} {R(s)}.
\end{equation}

\item
The new $(n+1)$-th trial is performed at the point
\begin{equation}\label{eq:22}
t^{n+1} = \frac{t_p-t_{p-1}}{2} - sign(v_p-v_{p-1}) \left( \frac{|v_p-v_{p-1}|}{M} \right)^N \frac{1}{2r} 
\end{equation}
the value $v^{n+1} = \varphi(t^{n+1})$ is computed and the iteration number $n$ is increased by 1.

\end{enumerate}

The method ALR described above is a modification of the base information-statistical algorithm \cite{Strongin2000,Sergeyev2013} for minimization functions satisfying the H{\"o}lder condition. The modification consists in introducing new characteristics (\ref{eq:17}) after initial stage of the search to improve the method's speedup on account of increasing values of characteristics for intervals containing current record (\ref{eq:18}).

Like its prototype, ARL can apply as the stopping condition the inequality
\begin{equation}\label{eq:23}
(t_p-t_{p-1})^{1/N} < \varepsilon,
\end{equation}
i.e., completes the search when the length of the interval with maximal characteristic (\ref{eq:21}) becomes less than the predefined accuracy $\varepsilon > 0$.

The sufficient condition of convergence to global minima in the problem (\ref{problem_fi}) with the objective function under H{\"o}lder property with constant $H$ is
\begin{equation}\label{eq:24}
m > 2^{2 - 1 / N}H.
\end{equation}

If $\varphi(t) = f(x(t)) $ where $x(t)$ is the Peano-type curve and $f(x)$ meets (\ref{lip_1}), then $H=2L\sqrt{N+3}$ and the sufficient condition takes the form 
\begin{equation}\label{eq:25}
m > 2^{3 - 1 / N}L\sqrt{N+3}.
\end{equation}

Finalizing the proposed general scheme  let us describe a procedure  of successive solving the problems of the family $S$ being scalar problems (\ref{conv_problem}) for different weight vectors $\lambda = (\lambda_1, \dots,\lambda_k)$ (\ref{lambda}).
